#Reproducible Research: Peer Assesment 1
##Author: Lavita Singhania

*Loading and preprocessing the data*

```{r reading data}
activity_data<-read.csv("activity.csv")
head(activity_data)
```

*Installing and calling ggplot for better visualization*
```{r loading packages, echo=FALSE}
library(ggplot2)
```

*Dividing and combining the data on daily basis*
```{r seggregating & combining data}
activity_data$date <- as.Date(activity_data$date, "%Y-%m-%d")
x<-aggregate(steps~date,data=activity_data,FUN=sum,na.rm=TRUE)
head(x)
```

*What is the mean total number of steps taken per day?*
```{r Question 1A}
qplot(steps,data=x,col="red",main="Total Number of steps each day",xlab="Steps")
```


The mean and median is: 
```{r Question 1B}
mean(x$steps,na.rm=TRUE)
median(x$steps,na.rm=TRUE)
```
 

*What is the average daily activity pattern?*
```{r Question 2A}
time_series <- tapply(activity_data$steps, activity_data$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Average number of steps taken", 
     col = "red")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r Question 2B}
max_interval <- which.max(time_series)
names(max_interval)
```

*Inputing missing values*

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r Question 3A}
activity_NA <- sum(is.na(activity_data))
activity_NA
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval,etc

First Na replaced by mean in 5 min interval
```{r Question 3B}
StepsAverage <- aggregate(steps ~ interval, data = activity_data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity_data)) {
    obs <- activity_data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```

Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r Question 3C}
new_activity <- activity_data
new_activity$steps <- fillNA
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r Question 3D}
StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
qplot(steps, data=StepsTotal2, main = "Total steps by day", xlab = "day", col="blue")
```

And the mean and median is
```{r Question 3E}
mean(StepsTotal2$steps)
median(StepsTotal2$steps)
```

**The mean is the same, there is a change in the median**

*Are there any differences in activity patterns between weekdays and weekends?*

For this part the weekdays() function may be of some help.
Create a new factor variable in the dataset with two levels weekday and weekend indicating whether a given date is a weekday or weekend day.
```{r Question 4A}
day <- weekdays(activity_data$date)
daylevel <- vector()
for (i in 1:nrow(activity_data)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity_data$daylevel <- daylevel
activity_data$daylevel <- factor(activity_data$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity_data, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:
```{r Question 4B}
qplot(x=interval, y=steps, data=stepsByDay, facets=.~daylevel, col="blue", geom="line", 
    xlab = "Interval", ylab = "Number of steps")
```




